State-level aggregation
========================================================

This vignette demonstrates the geographical aggregation of 
point data, to create choropleth maps summarising twitter data.

## Load the state data

```{r}
library(rgdal)
states <- readOGR("data/", "states")
summary(states)
plot(states)
head(states$STATE_NAME)
states@data

states <- spTransform(states, CRS("+init=epsg:4326"))
states <- states[-which(grepl("Alask|Haw", as.character(states$STATE_NAME))),]
tweets <- read.csv("1pSample.txt")
bbNY <- bbox(states[grepl("New York", as.character(states$STATE_NAME)),])
plot(states[grepl("New York", as.character(states$STATE_NAME)),])
plot(states)
plot(states, xlim=bbNY[1,], ylim=bbNY[2,])
points(bbNY, pch = 3)
plot(states[grepl("New York", as.character(states$STATE_NAME)),], col = "green", add = T)
points(tweets$lon, tweets$lat, col="blue")
```

Plotting with ggmap

```{r}
library(ggmap)
ggmap(get_map("Leeds"))
qmap(location="New York") +
geom_point(aes(tweets$lon, tweets$lat))

ggsave("ny.png")
```


Convert the tweets into a spatial (S4) class

```{r}
tweets <- SpatialPointsDataFrame(coords=matrix(c(tweets$lon, tweets$lat), ncol=2),data=tweets, proj4string=CRS("+init=epsg:4326"))
```

Now lets aggregate by number of tweets.

```{r}
statesAg1 <- aggregate(tweets["X"], states, length)
```

Aggregating by average n. friends in territory:

```{r}
statesAg2 <- aggregate(tweets["actor.friendsCount"], by = states, mean)
statesAg1$friends <- statesAg2$actor.friendsCount
statesAg1$id <- as.character(states$STATE_NAME)
```

## Visualisation

Preparing the data to plot with ggplot2.

```{r}
library(ggplot2)
library(dplyr)
sf <- fortify(statesAg1, region="id")
head(sf)
head(statesAg1@data)
sf <- inner_join(sf, statesAg1@data, by="id")
head(sf)
```

Plot.

```{r}
ggplot(sf, aes(long, lat, fill=X, group=group)) + geom_polygon() +
  scale_fill_gradient(low = "green", high = "red") + coord_map()

ggplot(sf, aes(long, lat, fill=friends, group=group)) + geom_polygon() +
  scale_fill_gradient(low = "blue", high = "orange") + coord_map()
```



